Reproducible Validation and Replication Studies in Nanoscale Physics

### 2020-07-27 tex/background.tex
small fix: background

In more recent work, we expanded the capabilities of PyGBE to applications of computational nanoplasmonics for biosensing \cite{ClementiETal2019}. 
Applying the long-wavelength limit, one can model the phenomenon of localized surface plasmon resonance (LSPR) via electrostatics, instead of the full Maxwell equations. 
This phenomenon is harnessed in nanosensors for detecting with high sensitivity the presence of biomolecules through shifts in resonance frequency. 
In LSPR, an electromagnetic wave excites the free electrons on the surface of a metallic nanoparticle. 
The name given to these electron-cloud vibrations is plasmons. 
In this setting, they resonate with the incoming field, and the energy is either absorbed by the nanoparticle or scattered in all directions, causing what's called extinction.
Since the resonance frequency depends on the dielectric environment, the change produced by a biomolecule approaching the metallic nanoparticle results in a frequency shift, and a means of detecting its presence.
The electrostatic approximation allows using the methods implemented in PyGBe, after substantial code development to permit using complex-valued permittivities, and to incorporate an external electric field into the model. 
These changes included re-writing the Krylov iterative solver to work with complex numbers, and splitting the treecode evaluation into real and imaginary parts. 
New functions were added to read from configuration files the data describing the electric field, to compute the dipole moment, and to compute the final quantity of interest: extinction cross-section. 
We reported a major new release of the software in 2017 \cite{ClementiETal2017}, and later  presented the mathematical formulation for electromagnetic scattering in the long-wavelength setting and its associated continuum and discretized integral equations, and results including verification activities and a sensitivity calculations on a biosensor model \cite{ClementiETal2019}. 
For verification, we conducted grid-convergence studies in two settings.
In the first, we set up a computational model for a spherical silver nanoparticle in a constant electric field, leading to an analytical solution for the extinction cross-section. 
The second case does not have a closed form solution: a spherical nanoparticle with a nearby protein (analyte), under an electric field. 
To estimate relative errors in the extinction cross-section, in this case, we made use of the method of Richardson extrapolation.
These verification activities build our confidence on the computational model---moreover, our work was published following careful reproducibility practices, including the release of reproducibility packages for all results. 
Nevertheless, the gold standard of confidence in the predictive capability of the model comes from validation: hunting for an opportunity to conduct validation studies with PyGBe in these settings led us through the twisted path that we report in this paper.

>>>>

In more recent work, we expanded the capabilities of PyGBE to applications of computational nanoplasmonics for biosensing \cite{ClementiETal2019}. 
Applying the long-wavelength limit, one can model the phenomenon of localized surface plasmon resonance (LSPR) via electrostatics, instead of the full Maxwell equations. 
This phenomenon is harnessed in nanosensors for detecting with high sensitivity the presence of biomolecules through shifts in resonance frequency. 
In LSPR, an electromagnetic wave excites the free electrons on the surface of a metallic nanoparticle. 
The name given to these electron-cloud vibrations is plasmons. 
In this setting, they resonate with the incoming field, and the energy is either absorbed by the nanoparticle or scattered in all directions, causing what's called extinction.
Since the resonance frequency depends on the dielectric environment, the change produced by a biomolecule approaching the metallic nanoparticle results in a frequency shift, and a means of detecting its presence.
The electrostatic approximation allows using the methods implemented in PyGBe, after substantial code development to permit using complex-valued permittivities, and to incorporate an external electric field into the model. 
These changes included re-writing the Krylov iterative solver to work with complex numbers, and splitting the treecode evaluation into real and imaginary parts. 
New functions were added to read from configuration files the data describing the electric field, to compute the dipole moment, and to compute the final quantity of interest: extinction cross-section. 
We reported a major new release of the software in 2017 \cite{ClementiETal2017}, and later  presented the mathematical formulation for electromagnetic scattering in the long-wavelength setting and its associated continuum and discretized integral equations, and results including verification activities and sensitivity calculations on a biosensor model \cite{ClementiETal2019}. 
For verification, we conducted grid-convergence studies in two settings.
In the first, we set up a computational model for a spherical silver nanoparticle in a constant electric field, leading to an analytical solution for the extinction cross-section. 
The second case does not have a closed form solution: a spherical nanoparticle with a nearby protein (analyte), under an electric field. 
To estimate relative errors in the extinction cross-section, in this case, we made use of the method of Richardson extrapolation.
These verification activities build our confidence on the computational model---moreover, our work was published following careful reproducibility practices, including the release of reproducibility packages for all results. 
Nevertheless, the gold standard of confidence in the predictive capability of the computational model comes from validation: our quest for an opportunity to conduct validation studies with PyGBe in these settings led us through the twisted path that we report in this paper.
